Concurrent Programming with CXL

### 2020-06-25 book/paper.tex
Added some more index terms

The producer/consumer pattern is common.  Processes may be arranged
in \emph{pipelines},
\index{pipeline}
where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers executing them in parallel.

@@

\section*{Exercises}
\begin{problems}
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\item Bonus question: implement a version of parallel qsort
(see Example~\ref{ex:qsort}) using the actor model.
\end{problems}

\chapter{Networking: Alternating Bit Protocol}
\label{ch:abp}
\index{alternating bit protocol}

@@

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account,
including failures of processes and failures of shared memory.
In this chapter, we will consider two actors, Alice and Bob.
Alice wants to send a sequence of application messages to Bob,
but the underlying network may lose messages.
The network does not re-order messages: when sending messages
$m_1$ and $m_2$ in that order, then if both messages are received,
$m_1$ is received before $m_2$.
Also, the network does not create messages out of nothing: if
message $m$ is received, then message $m$ was sent.

@@

>>>>

The producer/consumer pattern is common.  Processes may be arranged
in \emph{pipelines},
\index{pipeline}
where each upstream process is a producer and each downstream
process is a consumer.
Or processes may be arranged in a manager/worker pattern, with a manager
producing jobs and workers executing them in parallel.
Or, in the client/server model,
\index{client/server model}
some process may act as a \emph{server} that clients can send requests to
and receive responses from.

@@

\section*{Exercises}
\begin{problems}
\item \label{ex:cltsvr} A popular model is the \emph{client/server} model.
Here a single actor implements some service, like computing the square of a
number.
A client can send a message containing an integer to the server, and the
server returns a response message containing the square of that message.
Implement this.
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\end{problems}

\chapter{Networking: Alternating Bit Protocol}
\label{ch:abp}
\index{alternating bit protocol}
\index{protocol}
\index{distributed system}

@@

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures}
\index{failure}
into account,
including failures of processes and failures of shared memory.
In this chapter, we will consider two actors, Alice and Bob.
Alice wants to send a sequence of application messages to Bob,
but the underlying network may lose messages.
\index{network}
The network does not re-order messages: when sending messages
$m_1$ and $m_2$ in that order, then if both messages are received,
$m_1$ is received before $m_2$.
Also, the network does not create messages out of nothing: if
message $m$ is received, then message $m$ was sent.

@@

\section*{Exercises}
\begin{problems}
\item Exercise~\ref{ex:cltsvr} explored the \emph{client/server}.  It is popular
in distributed systems as well.
Develop a protocol for a single client and server using the same network
model as for the ABP protocol.
\end{problems}
